---
title: "p8105_hw5_ep3045"
author: "Ellen Park"
date: "2022-11-11"
output: github_document
---

```{r, include = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(purrr) 
library(forcats)

knitr::opts_chunk$set(collapse = TRUE, message = FALSE, dpi = 300, fig.width = 7)

```

# Problem 2 - Homicide Data

```{r}
homicide = 
  read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") 
```

The raw data contains `r nrow(homicide)` rows of homicide observations and `r ncol(homicide)` variables: homicide ID, date, the victim's first and last name, age, sex, city, state, latitude, longitude, and disposition. 

Dataframe with the total number of homicides and the number of unsolved homicides in each city:

```{r}
city = 
  homicide %>%
  mutate(
    city_state = paste(city, state, sep = ", ")) %>%
   mutate(
    city_state = if_else(city_state == "Tulsa, AL", "Tulsa, OK", city_state))  %>%
  group_by(city_state) %>%
  summarize(
    homicides_total = n(),
    homicides_unresolved = sum(disposition != "Closed by arrest"))
```

The proportion (w/ confidence limits) of homicides that are unsolved in Baltimore, MD:

```{r}
baltimore = 
  city %>%
  filter(city_state == "Baltimore, MD") 

prop_unresolved =
  prop.test(
    x = pull(baltimore, homicides_unresolved), 
    n = pull(baltimore, homicides_total))

prop_unresolved %>%
  broom::tidy() %>%
  select(estimate, conf.low, conf.high) %>%
  knitr::kable(
    col.names = c('Estimated Proportion', 'Lower Confidence Limit', 'Upper Confidence Limit'))
```

The proportion (with confidence limits) of homicides that are unsolved in all cities:

```{r}
city_nest = 
  city %>% 
  nest(data = homicides_total:homicides_unresolved)


prop_conflimits = function(df) {
  
  prop_unresolved =
  prop.test(
    x = pull(df, homicides_unresolved), 
    n = pull(df, homicides_total))

prop_unresolved %>%
  broom::tidy() %>%
  select(estimate, conf.low, conf.high)
}


city_nest %>% 
  mutate(prop_estimates = map(data, prop_conflimits)) %>% 
  unnest(data, prop_estimates) %>% 
  rename(
    prop_unresolved_est = estimate,
    conf_lower = conf.low,
    conf_upper = conf.high) -> city_prop
```

Plot showing the proportion (with confidence limits) of homicides that are unsolved in each city

```{r}
city_prop %>%
  mutate(city_state = fct_reorder(city_state, prop_unresolved_est)) %>%
  ggplot(aes(x = city_state, y = prop_unresolved_est)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_lower, ymax = conf_upper), width = 0.2) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Proportion of Unsolved Homicides in U.S. Cities",
    x = "City",
    y = "Proportion")
```